---
title: "LAB01"
author: "Group7"
output:
  html_document:
    df_print: paged
---

## Lab Exercize 1

#### This is a lab exercise

#### We will work through the following dataset in groups

#### We begin together with the dataset and preprocessing

#### (Optional) Now take a few minutes now and form groups with a maximum of three

#### After these data pre-prosessing steps, take some time and come up with the next step and be prepared to discuss your progress in 30 minutes

#### Write out the steps toward suggesting a "first try" predictive model, such as EDA plots etc.

#### You will submit your results as an individual on Bb

#### The R Summary of an extensive analysis is at the end of this section

```{r}
rm(list=ls())
```

## Load uswages

```{r}
library("faraway")
data(uswages)
```

```{r}
names(uswages)
```

## Print a taste of the data.

```{r, message=FALSE}
head(uswages,15)
```

## Data Cleaning

## Unusual values.

```{r}
summary(uswages)
```

## Since negative exper is not possible, convert to missing.

```{r}
uswages$exper[uswages$exper < 0] <- NA
```

```{r}
summary(uswages)
```

## Deal with factors

```{r}
# convert race, smsa, and pt to factor variables
uswages$race <- factor(uswages$race)
levels(uswages$race) <- c("White","Black")

uswages$smsa <- factor(uswages$smsa)
levels(uswages$smsa) <- c("No","Yes")

uswages$pt <- factor(uswages$pt)
levels(uswages$pt) <- c("No","Yes")

```

## Or with the with function.


```{r}
with(uswages, 
     race <- factor(race),
     levels(race) <- c("White", "Black")
     )
```

### Convert dummy var to one variable-factor


```{r}
# create region, a factor variable based on the four regions ne, mw, so, we
uswages <- data.frame(uswages, 
                      region = 
                        1*uswages$ne +
                        2*uswages$mw +
                        3*uswages$so +
                        4*uswages$we)
uswages$region <- factor(uswages$region)
levels(uswages$region) <- c("ne","mw","so","we")
head(uswages,15)
```

### Remove the dummies.

```{r}
# delete the four regions ne, mw, so, we
uswages <- subset(uswages,select=-c(ne:we))
```

```{r}
summary(uswages)
```

#### Missing data

- Delete all the missing data

```{r}
# Take care of NAs
uswages <- na.omit(uswages)
```

#### See summary of finished product

### Have one more look at the data summary

```{r}
summary(uswages)
```


```{r, fig.height=4.5, fig.width=7, message=FALSE}
# Explore scatterplot matrix using the continuous predictors
library("GGally")
ggpairs(uswages, columns = c("educ", "exper", "wage"), alpha = I(1/5))
# - `educ` and `exper` have a stronger relationship than either have with `wage`
# - `wage` is skewed to the right
# - Conclusion: Start with a simple model using `educ` as a predictor
```


```{r,  fig.height=9, fig.width=9, message=FALSE}
# Explore scatterplot matrix using the continuous predictors plus one factor
library("GGally")
ggpairs(uswages, 
        columns = c("region" ,"educ", "exper", "wage"),
        lower = list(
          continuous = "smooth",
          combo = "facetdensity",
          mapping = aes(color = region),
          alpha=I(1/100)
          )
)
  
```


```{r, fig.height=6, fig.width=7, message=FALSE}
# The pattern in "wage vs exper" needs a closer look

library("ggplot2")

# This plot features loess (default) smoother-fits per region

qplot(exper, 
      log(wage), 
      data = uswages, 
      color = region, 
      alpha = I(1/10)
      ) +
  geom_smooth()
# loess: "Local Polynomial Regression Fitting"
```

```{r, fig.height=6, fig.width=7, message=FALSE}
# Let's explore fitting a "global" quadratic model per region using log(wage) as the outcome

library("ggplot2")
plot1 <- qplot(exper, 
      log(wage), 
      data = uswages, 
      color = region, 
      alpha = I(1/5)
      ) +
  geom_smooth(method = "lm", 
              formula = y~poly(x, degree=2),
              se=FALSE
              ) +
  ggtitle("Quadratic Regressions per Region using exper ")
plot1
```

```{r,fig.height=6, fig.width=9, message=FALSE}
# This is further exploration of the relation between wage and exper within the part time and full time workers using pt

plot3 <- plot1 + 
  facet_grid(. ~ pt) +
    ggtitle("Quadratic Regressions per (`region`, `pt`) using `exper`")

plot3
```

```{r, fig.height=5, fig.width=7, message=FALSE}
# Upon having a closer look at the kernal density plot of `educ`

qplot(educ, data = uswages, geom = "density")

# It appears there are distinct clusters that coincide with "highschool","college", and "graduate" levels of education

# Creating a factor variable out of `educ` with levels equal to "highschool","college", and "graduate"

uswages$education <- cut(uswages$educ,
                     breaks=c(-Inf, 15, 17, Inf),
                     labels=c("highschool","college","graduate"))

# A boxplot of `log(wage)` per education

qplot(education, 
      log(wage), 
      data = uswages, 
      geom = "boxplot")

```

```{r, fig.height=5, fig.width=5, message=FALSE}
# Quadratic Regressions per Region using `exper`
plot7 <- qplot(exper,
               log(wage),
               data = uswages, 
               color = region, 
               alpha = I(1/5)
               ) +
  geom_smooth(method = "lm", 
              formula = y~poly(x, degree=2),
              se=FALSE
              ) +
  ggtitle("Quadratic Regressions per Region using `exper`")

# Quadratic Regressions per Educ using `exper`
plot8 <- qplot(exper, 
               log(wage), 
               data = uswages, 
               color = education, 
               alpha = I(1/5)
               ) +
  geom_smooth(method = "lm", 
              formula = y~poly(x, degree=2),
              se=FALSE
              ) +
  ggtitle("Quadratic Regressions per Educ using `exper`")


library("gridExtra")
grid.arrange(plot7, plot8, nrow=2)
```

```{r,  fig.height=9, fig.width=9, message=FALSE}
# Quadratic Regressions `log(wages)` using `exper` per (`region`, `pt`)
plot17 <- plot7 + 
  facet_grid(. ~ pt) +
  ggtitle("Quadratic Regressions `log(wages)` using `exper` per (`region`, `pt`)")
# Quadratic Regressions `log(wages)` using `exper` per (`education`, `pt`)
plot18 <- plot8 + 
  facet_grid(. ~ pt) +
  ggtitle("Quadratic Regressions `log(wages)` using `exper` per (`education`, `pt`)")
grid.arrange(plot17, plot18, nrow=2)
```


